### 2015-04-06 cug_paper/03_analysis.tex
Paper:  More edits & proofreading.  Last of the /todo boxes removed.

For both test series, the daemon was configured to write each rank's data to a separate file and the test measured the application's perceived write performance as the write size increased.  Note the word 'perceived'.  What was actually measured was how long it took for each rank of the application to copy its data into the GPU's memory.   The GPU cards in each of Titan's compute nodes have 6GB of RAM, though in practice only a little over 5GB is actually available to the user.  This means that for the first test series, with 8 ranks per node, write sizes of up to 512 MB were small enough for all ranks' writes to fit into the GPU memory.  For the second test series, using 16 ranks per node, write sizes up to 256MB would fit.  For both test series, once the write size exceeded the available GPU memory, the ranks would have to wait until the daemon was able to drain data out of the GPU memory and into the filesystem.

@@

Figure \ref{fig:results_8_nobars} shows the remaining results of the first test series.  As stated earlier, this series was run with 8 ranks per node and with 1, 2 \& 4 write threads in the daemon.  The graph shows a number of interesting features.  The first and most obvious, is that multiple write threads decrease performance.  Exactly why this is so is unclear, but it appears that having multiple threads read from the GPU's memory interfered with the individual ranks' ability to write to the GPU's memory.

Concentrating on the single thread performance, it's clear that the application benefits from using the GPU memory out to the 512MB write size.  This makes sense since eight ranks each writing 512MB is a total of 4GB and that will fit into the GPU memory.  Even at 1GB, the application sees somewhat improved performance because there is enough GPU memory to hold significant fraction of the data to be written.  It's not until the 2GB write size that the write performance is dominated by the filesystem's throughput.

Also obvious from the graph is the fact that small write sizes are not particularly efficient.  For write sizes less than 4MB, copying data to the GPU memory is slower than writing to the Lustre cache.  This is not surprising given the results shown in Figure \ref{fig:transfer_bw}.

@@

Figure \ref{fig:results_16_nobars} shows that again, the per-rank bandwidth is about half that shown in Figure \ref{fig:results_8_nobars} because there are twice as many ranks.  It's also clear that the application ranks see good write performance up to the 256MB write size, which is the largest size that will entirely fit into the GPU's memory.  It's also clear that, as in the first test series, writes must be at least 4MB in order to get reasonable performance copying the data into the GPU's memory.

>>>>

For both test series, the daemon was configured to write each rank's data to a separate file and the test measured the application's perceived write performance as the write size increased.  Note the word 'perceived'.  What was actually measured was how long it took for each rank of the application to copy its data into GPU memory.   The GPU cards in each of Titan's compute nodes have 6GB of RAM, though in practice only a little over 5GB is actually available to the user.  This means that for the first test series, with 8 ranks per node, write sizes of up to 512 MB were small enough for all ranks' writes to fit into GPU memory.  For the second test series, using 16 ranks per node, write sizes up to 256MB would fit.  For both test series, once the write size exceeded the available GPU memory, the ranks would have to wait until the daemon was able to drain data out of GPU memory and into the filesystem.

@@

Figure \ref{fig:results_8_nobars} shows the remaining results of the first test series.  As stated earlier, this series was run with 8 ranks per node and with 1, 2 \& 4 write threads in the daemon.  The graph shows a number of interesting features.  The first and most obvious, is that multiple write threads decrease performance.  Exactly why this is so is unclear, but it appears that having multiple threads read from GPU memory interfered with the individual ranks' ability to copy data into GPU memory.

Concentrating on the single thread performance, it's clear that the application benefits from using GPU memory out to the 512MB write size.  This makes sense since eight ranks each writing 512MB is a total of 4GB and that will fit into the available GPU memory.  Even at 1GB, the application sees somewhat improved performance because there is enough GPU memory to hold significant fraction of the data to be written.  It's not until the 2GB write size that the write performance is dominated by the filesystem's throughput.

Also obvious from the graph is the fact that small write sizes are not particularly efficient.  For write sizes less than 4MB, copying data to GPU memory is slower than writing to the Lustre cache.  This is not surprising given the results shown in Figure \ref{fig:transfer_bw}.

@@

Figure \ref{fig:results_16_nobars} shows that again, the per-rank bandwidth is about half that shown in Figure \ref{fig:results_8_nobars} because there are twice as many ranks.  It's also clear that the application ranks see good write performance up to the 256MB write size, which is the largest size that will entirely fit into GPU memory.  It's also clear that, as in the first test series, writes must be at least 4MB in order to get reasonable performance copying the data into GPU memory.

### 2015-04-06 cug_paper/04_futurework.tex
Paper:  More edits & proofreading.  Last of the /todo boxes removed.

Another feature the authors would like to add is the ability to use regular system RAM in addition to the GPU RAM.  It is well known that some applications do not use all the RAM on the compute node.  It is debatable whether letting the daemon use the memory is better than simply letting the operating system use it as page cache.  However, both the Linux kernel page cache and the client-side cache have fairly low limits on the amount of dirty pages they will allow.  Some very basic initial testing implies that allowing the daemon to allocate memory may be more useful than letting the OS use the same memory for page cache, but further testing is required to be certain.

>>>>

Another feature the authors would like to add is the ability to use regular system RAM in addition to the GPU RAM.  It is well known that some applications do not use all the RAM on the compute node.  It is debatable whether letting the daemon use the memory is better than simply letting the operating system use it as  Lustre client cache or OS page cacher.  However, as was discussed in Section \ref{sec:analysis} and Section \ref{sec:conclusion}, the Lustre client-side cache has fairly low limits on the amount of dirty pages it will allow.  (The Linux kernel page cache has similar low limits on dirty pages.)  Some very basic initial testing implies that allowing the daemon to allocate memory may be more useful than letting the OS use the same memory for page cache, but further testing is required to be certain.

### 2015-04-06 cug_paper/05_conclusion.tex
Paper:  More edits & proofreading.  Last of the /todo boxes removed.

For the first test series (8 ranks/node), the results show a clear performance improvement compared to writing directly to the filesystem.   More data can be cached in the GPU's memory than in the Lustre client-side cache and the average write speed is, surprisingly, slightly higher.  It is, however, necessary to use relatively large writes.   Comparing Figure \ref{fig:results_base_8} and Figure \ref{fig:results_8_nobars},  it can be seen that write sizes must be at least 4MB before transfers to the GPU memory outperform writes to the Lustre cache.

@@

The results for the second test series (16 ranks/node) are less clear.  As with the first test series, write sizes need to be at least 4MB Looking closely at Figure \ref{fig:results_base_16} and Figure \ref{fig:results_16_nobars}, it appears that write sizes that can fit into the GPU's memory perform slightly better than just writing to disk.  However, write sizes that exceed the capacity of the GPU's memory are actually a little slower than writing straight to the filesystem.

There are several other complications to consider, though.
First, Since the test application did no actual computation, there were plenty of cores available for the daemon to execute on.  In a real HPC application, this wouldn't be the case and the application's performance would almost certainly be negatively affected by the context switches.  On the other hand, an application that is running 16 ranks/node is unlikely to leave much system memory free for use by the Lustre client-side cache.  So, the improvement from caching writes in the GPU's memory might offset the disruption caused by oversubscribing the cores.  Whether or not this technique provides a net performance improvement in such a configuration will probably depend on the particular application.

@@

In general, the results show that using the GPU's memory to cache filesystem writes provides approximately the same performance as the Lustre client-side cache.  The difference is that in order to benefit from the Lustre cache, the writes must be spread across a sufficient number of OST's and the memory must actually be available.  Whether or not these conditions can be met depends on the particular application.

>>>>

For the first test series (8 ranks/node), the results show a clear performance improvement compared to writing directly to the filesystem.   More data can be cached in GPU memory than in the Lustre client-side cache and the average write speed is, surprisingly, slightly higher.  It is, however, necessary to use relatively large writes.   Comparing Figure \ref{fig:results_base_8} and Figure \ref{fig:results_8_nobars},  it can be seen that write sizes must be at least 4MB before transfers to GPU memory outperform writes to the Lustre cache.

@@

The results for the second test series (16 ranks/node) are less clear.  As with the first test series, write sizes need to be at least 4MB Looking closely at Figure \ref{fig:results_base_16} and Figure \ref{fig:results_16_nobars}, it appears that write sizes that can fit into GPU memory perform slightly better than just writing to disk.  However, write sizes that exceed the capacity of the GPU's memory are actually a little slower than writing straight to the filesystem.

There are several other complications to consider, though.
First, Since the test application did no actual computation, there were plenty of cores available for the daemon to execute on.  In a real HPC application, this wouldn't be the case and the application's performance would almost certainly be negatively affected by the context switches.  On the other hand, an application that is running 16 ranks/node is unlikely to leave much system memory free for use by the Lustre client-side cache.  So, the improvement from caching writes in GPU memory might offset the disruption caused by oversubscribing the cores.  Whether or not this technique provides a net performance improvement in such a configuration will probably depend on the particular application.

@@

In general, the results show that using the GPU memory to cache filesystem writes provides approximately the same performance as the Lustre client-side cache.  The difference is that the Lustre cache uses system memory which may not actually be available.  Even if it is, the user must ensure that writes are spread across a sufficient number of OST's to make sue of it.  Neither of these conditions apply when using GPU memory as a cache.
